### 2024-11-20 paper/paper.qmd
initial rproj structure

---
title: "Predicting the fish hunting population"
subtitle: ""
author: 
  - Shreya Sakura Noskor
thanks: "Code and data are available at: [https://github.com/NotSakura/FisheriesData.git](https://github.com/NotSakura/FisheriesData.git)."
date: today
date-format: long
abstract: ""
format:
  pdf:
    bibliography: references.bib
    documentclass: article
    geometry: margin = 1in
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)
library(janitor)
library(lubridate)
library(broom)
library(modelsummary)
library(rstanarm)
library(knitr)
```

# Introduction

# Data {#sec-data}

## Overview

## Methodology and Measurement

## Data Visualization {#sec-graphs}

## Outcome and Predictor variables

# Model

## Model Step- Up

### Model justification

## Model Results

# Results

# Discussion

## Weaknesses and next steps

\newpage

\appendix

# Appendix 1

# Appendix 2

# References

### 2024-11-20 scripts/00-simulate_data.R

@@ -1,0 +1,42 @@
+#### Preamble ####
+# Purpose: Simulates a dataset of American election
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 3 November 2024
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
+# Pre-requisites: The `tidyverse` package must be installed
+
+
+
+#### Workspace setup ####
+library(tidyverse)
+# Load necessary library
+library(dplyr)
+
+# Set seed for reproducibility
+set.seed(123)
+
+# Parameters for the simulation
+num_polls <- 100
+pollsters <- c("YouGov", "New York Times", "Ipsos", "The Washington Post", "1892 Polling")
+states <- c("Arizona", "Georgia", "Wisconsin", "Nevada", "Pennsylvania", "North Carolina", "Michigan")
+candidates <- c("Donald Trump", "Kamala Harris", "Jill Stein", "Chase Oliver")
+
+# Generate simulated data
+poll_data <- data.frame(
+  pollster = sample(pollsters, num_polls, replace = TRUE),
+  sample_size = sample(c(800, 1000, 1200, 1500, 2000), num_polls, replace = TRUE),
+  state = sample(states, num_polls, replace = TRUE),
+  candidate_name = sample(candidates, num_polls, replace = TRUE),
+  pct = round(runif(num_polls, 40, 60), 1),  # Simulate percentages between 40 and 60
+  start_date = sample(seq(as.Date('2024-10-01'), as.Date('2024-10-10'), by="days"), num_polls, replace = TRUE),
+  end_date = sample(seq(as.Date('2024-10-11'), as.Date('2024-10-20'), by="days"), num_polls, replace = TRUE),
+  pollscore = round(runif(num_polls, -2.0, 0), 1),  # Simulate poll scores
+  numeric_grade = round(runif(num_polls, 2.0, 5.0), 1)  # Simulate numeric grades
+)
+
+# View the first few rows of the simulated data
+head(poll_data)
+
+
+#### Save data ####
+write_csv(poll_data, "data/00-simulated_data/simulated_data.csv")

### 2024-11-20 scripts/01-test_simulated_data.R

@@ -1,0 +1,55 @@
+#### Preamble ####
+# Purpose: Tests the structure and validity of the simulated American election data
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 3 November 2024
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
+# Pre-requisites: 
+  # - The `tidyverse` package must be installed and loaded
+  # - 00-simulate_data.R must have been run
+
+
+
+install.packages("validate")
+install.packages("testthat")
+
+#### Workspace setup ####
+library(tidyverse)
+
+# Load necessary libraries
+# Load necessary libraries
+library(testthat)
+library(validate)
+library(readr)
+
+# Load the simulated data
+poll_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+
+# Define validation rules
+rules <- validator(
+  n_polls = nrow(poll_data) == 100,
+  n_cols = ncol(poll_data) == 9,
+  is_character_pollster = all(sapply(poll_data$pollster, is.character)),
+  is_numeric_sample_size = all(sapply(poll_data$sample_size, is.numeric)),
+  is_character_state = all(sapply(poll_data$state, is.character)),
+  is_character_candidate = all(sapply(poll_data$candidate_name, is.character)),
+  is_numeric_pct = all(sapply(poll_data$pct, is.numeric)),
+  is_date_start = all(sapply(poll_data$start_date, is.Date)),
+  is_date_end = all(sapply(poll_data$end_date, is.Date)),
+  is_numeric_pollscore = all(sapply(poll_data$pollscore, is.numeric)),
+  is_numeric_grade = all(sapply(poll_data$numeric_grade, is.numeric)),
+  pollster_in_set = all(poll_data$pollster %in% c("YouGov", "New York Times", "Ipsos", "The Washington Post", "1892 Polling")),
+  sample_size_in_set = all(poll_data$sample_size %in% c(800, 1000, 1200, 1500, 2000)),
+  state_in_set = all(poll_data$state %in% c("Arizona", "Georgia", "Wisconsin", "Nevada", "Pennsylvania", "North Carolina", "Michigan")),
+  candidate_in_set = all(poll_data$candidate_name %in% c("Donald Trump", "Kamala Harris", "Jill Stein", "Chase Oliver")),
[…]

### 2024-11-20 scripts/02-download_data.R

@@ -1,0 +1,26 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from [data](https://projects.fivethirtyeight.com/polls/president-general/2024/national/)
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 3 November 2024
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
+# Pre-requisites: go to the site above and search for “Download the data”, then select Presidential general election polls (current cycle), then “Download”
+
+
+
+library(tidyverse)
+
+# Load necessary libraries
+library(testthat)
+library(validate)
+library(readr)
+
+
+# Load the simulated data
+#the data is already downloaded in the repository. 
+data <- read.csv("../data/01-raw_data/raw_data.csv")
+
+#### Save data ####
+
+
+
+         

### 2024-11-20 scripts/03-clean_data.R

@@ -1,0 +1,27 @@
+#### Preamble ####
+# Purpose: Cleans the raw election data by filtering and deleting rows
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 3 November 2024
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
+
+#### Workspace setup ####
+library(tidyverse)
+library(arrow)
+library(janitor)
+
+
+#### Clean data ####
+data <- read.csv("./data/01-raw_data/raw_data.csv") |>
+  clean_names()
+
+preddata <- data %>%
+  filter(numeric_grade >1.5 & transparency_score >6) %>%
+  select(pollster, sample_size, state, candidate_name, pct, start_date, end_date, pollscore, numeric_grade) %>%
+  filter(!is.na(pollscore)) %>%
+  filter(!is.na(state) & state != "")
+
+
+
+#### Save data ####
+write_csv(preddata, "./data/02-analysis_data/cleaned.csv")
+arrow::write_parquet(preddata, "./data/02-analysis_data/cleaned.parquet")

### 2024-11-20 scripts/04-test_analysis_data.R

@@ -1,0 +1,47 @@
+#### Preamble ####
+# Purpose: Tests the data set for any out of place values
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 3 November 2024
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+# Load necessary libraries
+# Load necessary libraries
+library(testthat)
+library(validate)
+library(readr)
+
+# Load the simulated data
+poll_data <- read.csv("./data/01-raw_data/raw_data.csv")
+
+# Define validation rules
+rules <- validator(
+  n_polls = nrow(poll_data) == 100,
+  n_cols = ncol(poll_data) == 9,
+  is_character_pollster = all(sapply(poll_data$pollster, is.character)),
+  is_numeric_sample_size = all(sapply(poll_data$sample_size, is.numeric)),
+  is_character_state = all(sapply(poll_data$state, is.character)),
+  is_character_candidate = all(sapply(poll_data$candidate_name, is.character)),
+  is_numeric_pct = all(sapply(poll_data$pct, is.numeric)),
+  is_date_start = all(sapply(poll_data$start_date, is.Date)),
+  is_date_end = all(sapply(poll_data$end_date, is.Date)),
+  is_numeric_pollscore = all(sapply(poll_data$pollscore, is.numeric)),
+  is_numeric_grade = all(sapply(poll_data$numeric_grade, is.numeric)),
+  pollster_in_set = all(poll_data$pollster %in% c("YouGov", "New York Times", "Ipsos", "The Washington Post", "1892 Polling")),
+  sample_size_in_set = all(poll_data$sample_size %in% c(800, 1000, 1200, 1500, 2000)),
+  state_in_set = all(poll_data$state %in% c("Arizona", "Georgia", "Wisconsin", "Nevada", "Pennsylvania", "North Carolina", "Michigan")),
+  candidate_in_set = all(poll_data$candidate_name %in% c("Donald Trump", "Kamala Harris", "Jill Stein", "Chase Oliver")),
+  pct_range = all(poll_data$pct >= 40 & poll_data$pct <= 60),
+  start_before_end = all(poll_data$start_date < poll_data$end_date),
+  pollscore_negative = all(poll_data$pollscore < 0),
[…]

### 2024-11-20 scripts/05-exploratory_data_analysis.R

@@ -1,0 +1,54 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 3 November 2024
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+
+#### Read data ####
+preddata< - read_csv(preddata, "./data/02-analysis_data/cleaned.csv")
+
+harrisdata <- preddata %>%
+  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
+  filter(candidate_name == "Kamala Harris") %>%
+  mutate(end_date = mdy(end_date)) %>%
+  filter(end_date >= as.Date("2024-07-21")) %>%
+  mutate(num_harris = round((pct / 100) * sample_size, 0))
+
+### Model data ####
+harrisdata <- harrisdata %>%
+  mutate(
+    end_date_num = as.numeric(end_date - min(end_date)))
+
+harrisdata <- harrisdata |>
+  mutate(state = factor(state))
+
+# Define Model
+model <- cbind(num_harris, sample_size - num_harris) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)
+
+priors <- normal(0, 2.5, autoscale = TRUE)
+
+bayesian_model <- stan_glmer(
+  formula = model,
+  data = harrisdata,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  weights = harrisdata$numeric_grade,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95)
+
+
+
+#### Save model ####
+saveRDS(
+  bayesian_model,
+  file = "models/bay_model.rds"
+)
+
+

### 2024-11-20 scripts/06-model_data.R

@@ -1,0 +1,58 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 3 November 2024
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+
+#### Read data ####
+preddata< - read_csv(preddata, "./data/02-analysis_data/cleaned.csv")
+
+trumpdata <- preddata %>%
+  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
+  filter(candidate_name == "Donald Trump") %>%
+  mutate(end_date = mdy(end_date)) %>%
+  filter(end_date >= as.Date("2024-07-21")) %>%
+  mutate(num_trump = round((pct / 100) * sample_size, 0))
+
+ggplot(trumpdata, aes(x = reorder(state, pct, median), y = pct, fill = state)) +
+  geom_boxplot() +
+  coord_flip() +  # Flip for better readability
+  labs(x = "State", y = "Support Percentage") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+trumpdata <- trumpdata %>%
+  mutate(
+    end_date_num = as.numeric(end_date - min(end_date)))
+
+trumpdata <- trumpdata |>
+  mutate(state = factor(state))
+
+# Define Model
+model2 <- cbind(num_trump, sample_size - num_trump) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)
+
+priors <- normal(0, 2.5, autoscale = TRUE)
+
+bayesian_model2 <- stan_glmer(
+  formula = model2,
+  data = trumpdata,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  weights = trumpdata$numeric_grade,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95)
+
+
+#### Save model ####
+saveRDS(
+  bayesian_model2,
+  file = "models/bayesian_model2.rds"
+)
+
+

### 2024-11-26 scripts/00-simulate_data.R
updated scripts; all are done except model

@@ -1,8 +1,8 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of American election
-# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
-# Date: 3 November 2024
-# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
+# Purpose: Simulates a dataset of Fishing data
+# Author: Sakura Noskor
+# Date: 23 November 2024
+# Contact: sakura.noskor@mail.utoronto.ca
 # Pre-requisites: The `tidyverse` package must be installed
 
 
@@ -11,32 +11,32 @@
 library(tidyverse)
 # Load necessary library
 library(dplyr)
+set.seed(835)  # For reproducibility
 
-# Set seed for reproducibility
-set.seed(123)
 
-# Parameters for the simulation
-num_polls <- 100
-pollsters <- c("YouGov", "New York Times", "Ipsos", "The Washington Post", "1892 Polling")
-states <- c("Arizona", "Georgia", "Wisconsin", "Nevada", "Pennsylvania", "North Carolina", "Michigan")
-candidates <- c("Donald Trump", "Kamala Harris", "Jill Stein", "Chase Oliver")
+species <- c("Salmon", "Tuna", "Cod", "Herring", "Mackerel")
+areas <- c("Atlantic", "Pacific", "Indian Ocean", "Mediterranean", "Arctic")
+countries <- c("USA", "Canada", "Norway", "Japan", "Australia")
 
-# Generate simulated data
-poll_data <- data.frame(
-  pollster = sample(pollsters, num_polls, replace = TRUE),
-  sample_size = sample(c(800, 1000, 1200, 1500, 2000), num_polls, replace = TRUE),
-  state = sample(states, num_polls, replace = TRUE),
-  candidate_name = sample(candidates, num_polls, replace = TRUE),
-  pct = round(runif(num_polls, 40, 60), 1),  # Simulate percentages between 40 and 60
-  start_date = sample(seq(as.Date('2024-10-01'), as.Date('2024-10-10'), by="days"), num_polls, replace = TRUE),
-  end_date = sample(seq(as.Date('2024-10-11'), as.Date('2024-10-20'), by="days"), num_polls, replace = TRUE),
-  pollscore = round(runif(num_polls, -2.0, 0), 1),  # Simulate poll scores
[…]

### 2024-11-26 scripts/01-test_simulated_data.R

@@ -1,16 +1,17 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated American election data
-# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
-# Date: 3 November 2024
-# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
+# Purpose: tests the simuated dataset of Fishing data
+# Author: Sakura Noskor
+# Date: 23 November 2024
+# Contact: sakura.noskor@mail.utoronto.ca
 # Pre-requisites: 
   # - The `tidyverse` package must be installed and loaded
   # - 00-simulate_data.R must have been run
+  # - necessary packages must be installed and loaded
 
 
 
-install.packages("validate")
-install.packages("testthat")
+#install.packages("validate")
+#install.packages("validate")
 
 #### Workspace setup ####
 library(tidyverse)
@@ -22,34 +23,60 @@ library(validate)
 library(readr)
 
 # Load the simulated data
-poll_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+fish_catch_data <- read_csv("data/00-simulated_data/simulated_data.csv")
 
-# Define validation rules
-rules <- validator(
-  n_polls = nrow(poll_data) == 100,
-  n_cols = ncol(poll_data) == 9,
-  is_character_pollster = all(sapply(poll_data$pollster, is.character)),
-  is_numeric_sample_size = all(sapply(poll_data$sample_size, is.numeric)),
-  is_character_state = all(sapply(poll_data$state, is.character)),
-  is_character_candidate = all(sapply(poll_data$candidate_name, is.character)),
-  is_numeric_pct = all(sapply(poll_data$pct, is.numeric)),
-  is_date_start = all(sapply(poll_data$start_date, is.Date)),
-  is_date_end = all(sapply(poll_data$end_date, is.Date)),
-  is_numeric_pollscore = all(sapply(poll_data$pollscore, is.numeric)),
-  is_numeric_grade = all(sapply(poll_data$numeric_grade, is.numeric)),
-  pollster_in_set = all(poll_data$pollster %in% c("YouGov", "New York Times", "Ipsos", "The Washington Post", "1892 Polling")),
-  sample_size_in_set = all(poll_data$sample_size %in% c(800, 1000, 1200, 1500, 2000)),
[…]

### 2024-11-26 scripts/02-download_data.R

@@ -1,25 +1,17 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [data](https://projects.fivethirtyeight.com/polls/president-general/2024/national/)
-# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
-# Date: 3 November 2024
-# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
-# Pre-requisites: go to the site above and search for “Download the data”, then select Presidential general election polls (current cycle), then “Download”
-
-
+# Purpose: Simulates a dataset of Fishing data
+# Author: Sakura Noskor
+# Date: 23 November 2024
+# Contact: sakura.noskor@mail.utoronto.ca
+# Pre-requisites: go to "https://www.npafc.org/statistics/" and look for "NPAFC Catch Statistics (updated 28 June 2024)". Click on that you will get the csv file containing all the data. 
 
 library(tidyverse)
-
-# Load necessary libraries
-library(testthat)
-library(validate)
-library(readr)
 
 
 # Load the simulated data
 #the data is already downloaded in the repository. 
-data <- read.csv("../data/01-raw_data/raw_data.csv")
+data <- read.xlsx('../data/01-raw_data/NPAFC_Catch_Stat-1925-2023.xlsx', sheet = 1) 
 
-#### Save data ####
 
 
 

### 2024-11-26 scripts/03-clean_data.R

@@ -1,27 +1,44 @@
 #### Preamble ####
-# Purpose: Cleans the raw election data by filtering and deleting rows
-# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
-# Date: 3 November 2024
-# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
+# Purpose: Cleans fishing raw dataset
+# Author: Sakura Noskor
+# Date: 23 November 2024
+# Contact: sakura.noskor@mail.utoronto.ca
+
+#install.packages("readxl")
 
 #### Workspace setup ####
 library(tidyverse)
 library(arrow)
 library(janitor)
+library(readxl)
 
 
 #### Clean data ####
-data <- read.csv("./data/01-raw_data/raw_data.csv") |>
-  clean_names()
 
-preddata <- data %>%
-  filter(numeric_grade >1.5 & transparency_score >6) %>%
-  select(pollster, sample_size, state, candidate_name, pct, start_date, end_date, pollscore, numeric_grade) %>%
-  filter(!is.na(pollscore)) %>%
-  filter(!is.na(state) & state != "")
+data <- read.xlsx('../data/01-raw_data/NPAFC_Catch_Stat-1925-2023.xlsx', sheet = 1) 
+print(data)
+colnames(data) <- data[1, ]
+data <- data[-1, ]
 
+cleaned_pac2 <- data %>%
+  filter(`Data Type` == "Number (000's)", `Whole Country/Province/State` == "Whole country", Species == "Total")  %>%
+  select(Country, Species, `Catch Type`, `2022`) %>%
+  filter(!is.na(`2022`) & `2022` != 0)
+cleaned_pac2
+
+
+cleaned_pac <- data %>%
+  filter(Species == "Total", `Data Type` == "Number (000's)", `Reporting Area` == "Whole country") %>%
+  filter(!is.na(Country)) %>%
+  filter(!is.na(`1925`) & `1925` != 0) %>%
+  filter(!is.na(`1946`) & `1946` != 0)
+
+cleaned_pac
 
 
 #### Save data ####
-write_csv(preddata, "./data/02-analysis_data/cleaned.csv")
-arrow::write_parquet(preddata, "./data/02-analysis_data/cleaned.parquet")
[…]

### 2024-11-26 scripts/04-test_analysis_data.R

@@ -1,9 +1,17 @@
 #### Preamble ####
-# Purpose: Tests the data set for any out of place values
-# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
-# Date: 3 November 2024
-# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
+# Purpose: tests the dataset of Fishing data
+# Author: Sakura Noskor
+# Date: 23 November 2024
+# Contact: sakura.noskor@mail.utoronto.ca
+# Pre-requisites: 
+# - The `tidyverse` package must be installed and loaded
+# - 00-simulate_data.R must have been run
+# - necessary packages must be installed and loaded
 
+
+
+#install.packages("validate")
+#install.packages("validate")
 
 #### Workspace setup ####
 library(tidyverse)
@@ -15,33 +23,60 @@ library(validate)
 library(readr)
 
 # Load the simulated data
-poll_data <- read.csv("./data/01-raw_data/raw_data.csv")
+fish_catch_data <- read.xlsx('../data/01-raw_data/NPAFC_Catch_Stat-1925-2023.xlsx', sheet = 1) 
 
-# Define validation rules
-rules <- validator(
-  n_polls = nrow(poll_data) == 100,
-  n_cols = ncol(poll_data) == 9,
-  is_character_pollster = all(sapply(poll_data$pollster, is.character)),
-  is_numeric_sample_size = all(sapply(poll_data$sample_size, is.numeric)),
-  is_character_state = all(sapply(poll_data$state, is.character)),
-  is_character_candidate = all(sapply(poll_data$candidate_name, is.character)),
-  is_numeric_pct = all(sapply(poll_data$pct, is.numeric)),
-  is_date_start = all(sapply(poll_data$start_date, is.Date)),
-  is_date_end = all(sapply(poll_data$end_date, is.Date)),
-  is_numeric_pollscore = all(sapply(poll_data$pollscore, is.numeric)),
-  is_numeric_grade = all(sapply(poll_data$numeric_grade, is.numeric)),
-  pollster_in_set = all(poll_data$pollster %in% c("YouGov", "New York Times", "Ipsos", "The Washington Post", "1892 Polling")),
-  sample_size_in_set = all(poll_data$sample_size %in% c(800, 1000, 1200, 1500, 2000)),
[…]

### 2024-11-26 scripts/03-clean_data.R
models scripts is done

@@ -11,6 +11,7 @@ library(tidyverse)
 library(arrow)
 library(janitor)
 library(readxl)
+library(openxlsx)
 
 
 #### Clean data ####

### 2024-11-26 scripts/06-model_data.R

@@ -1,57 +1,77 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
-# Date: 3 November 2024
-# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
+# Purpose: models the data so it predictis how much the fishing increases per year
+# Author: Sakura Noskor
+# Date: 23 November 2024
+# Contact: sakura.noskor@mail.utoronto.ca
 
 #### Workspace setup ####
-library(tidyverse)
+library(tidyr)
+library(dplyr)
+
+# Convert columns that should be years to numeric
+cleaned_pac <- read.csv("./data/02-analysis_data/cleaned_pacific_allyear.csv")
+
+temp <- cleaned_pac  %>%
+  select(-`Whole Country/Province/State`, -`Data Type`, -`Reporting Area`, -Species, -`Catch Type`)
+
+temp <- temp %>%
+  mutate(across(`1925`:`2023`, as.numeric))
+
+long_data <- temp %>%
+  pivot_longer(
+    cols = starts_with("19") | starts_with("20"),  # Select columns starting with "19" or "20"
+    names_to = "Year",
+    values_to = "Catch"
+  ) %>%
+  mutate(Year = as.numeric(Year))
+
 library(rstanarm)
 
-#### Read data ####
-preddata< - read_csv(preddata, "./data/02-analysis_data/cleaned.csv")
+model_formula <- Catch ~ Year + (1 | Country)
 
-trumpdata <- preddata %>%
-  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
-  filter(candidate_name == "Donald Trump") %>%
-  mutate(end_date = mdy(end_date)) %>%
-  filter(end_date >= as.Date("2024-07-21")) %>%
-  mutate(num_trump = round((pct / 100) * sample_size, 0))
-
-ggplot(trumpdata, aes(x = reorder(state, pct, median), y = pct, fill = state)) +
-  geom_boxplot() +
-  coord_flip() +  # Flip for better readability
-  labs(x = "State", y = "Support Percentage") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-trumpdata <- trumpdata %>%
[…]

### 2024-11-26 paper/paper.qmd
paper to commit
references and paper updated; still have to write the paragraphs and the graph for each of the variables i showed.

@@
 ---
-title: "Predicting the fish hunting population"
-subtitle: ""
+title: "Analysing the Fish Hunting Numbers in North Pacific Ocean"
+subtitle: "Using Baysian Modeling, to find that number of fish caught has increased"
 author: 
@@
 date-format: long
-abstract: ""
+abstract: "We analyzed data on fish catches in the North Pacific Ocean using Bayesian modeling.
+Our analysis shows that the number of fish caught has increased over time.
+This suggests that fishing activities in the region have intensified.
+Understanding this trend is important for managing fish populations and ensuring the sustainability of the ocean's resources.
+This is important as the Pacific is full of different kind of fish like, tuna and salmon that are essential in the food chain, but are being overhunted for food and recreational purposes."
 format:
@@
 library(tidyverse)
+library(janitor)
+library(readxl)
 library(dplyr)
-library(janitor)
 library(lubridate)
@@
 library(knitr)
+library(openxlsx)
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+data <- read.xlsx('../data/01-raw_data/NPAFC_Catch_Stat-1925-2023.xlsx', sheet = 1) 
+print(data)
+colnames(data) <- data[1, ]
+data <- data[-1, ]
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+cleaned_pac2 <- data %>%
+  filter(`Data Type` == "Number (000's)", `Whole Country/Province/State` == "Whole country", Species == "Total")  %>%
+  select(Country, Species, `Catch Type`, `2022`) %>%
+  filter(!is.na(`2022`) & `2022` != 0)
+
+cleaned_pac <- data %>%
+  filter(Species == "Total", `Data Type` == "Number (000's)", `Reporting Area` == "Whole country") %>%
+  filter(!is.na(Country)) %>%
+  filter(!is.na(`1925`) & `1925` != 0) %>%
+  filter(!is.na(`1946`) & `1946` != 0)
+
+koreadata <- data %>%
+  filter(Species == "Total", `Data Type` == "Number (000's)", `Reporting Area` == "Whole country", Country == "Korea") %>%
+  filter(!is.na(Country))
+
+combined_data <- bind_rows(cleaned_pac, koreadata)
+
 ```
@@
 # Data {#sec-data}
+
+China, Japan, Russia
 
@@
 
+we make the assumption that the coulms that say whole country it also include the provinces and different areas number as well.
+
 ## Outcome and Predictor variables
@@
 ## Model Step- Up
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+temp <- cleaned_pac  %>%
+  select(-`Whole Country/Province/State`, -`Data Type`, -`Reporting Area`, -Species, -`Catch Type`)
+temp
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+library(tidyr)
+library(dplyr)
+
+# Convert columns that should be years to numeric
+temp <- temp %>%
+  mutate(across(`1925`:`2023`, as.numeric))
+
+long_data <- temp %>%
+  pivot_longer(
+    cols = starts_with("19") | starts_with("20"),  # Select columns starting with "19" or "20"
+    names_to = "Year",
+    values_to = "Catch"
+  ) %>%
+  mutate(Year = as.numeric(Year))
+
+library(rstanarm)
+
+model_formula <- Catch ~ Year + (1 | Country)
+
+bayesian_model <- stan_glmer(
+  formula = model_formula,
+  data = long_data,
+  family = gaussian(),  # Adjust this if using a different distribution for your data
+  prior = normal(0, 2.5, autoscale = TRUE),
+  prior_intercept = normal(0, 2.5, autoscale = TRUE),
+  seed = 123,
+  adapt_delta = 0.95,
+  cores = 4
+)
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+modelsummary(bayesian_model)
+```
+
+Model 2
+
+```{r}
+
+library(brms)
+
+# Specify priors
+priors <- c(
+  set_prior("normal(0, 2.5)", class = "Intercept"),   # Prior for the intercept
+  set_prior("normal(0, 2.5)", class = "sd")           # Prior for the standard deviation of random effects (Country)
+)
+
+# Fit the model using brms
+bayesian_model_commercial_brms <- brm(
+  formula = Is_Commercial ~ (1 | Country),  # Random effect for Country
+  data = cleaned_pac2,
+  family = bernoulli(),                    # Bernoulli family for binary outcomes
+  prior = priors,                          # Use the prior argument here
+  seed = 123
+)
+
+# Check the model summary
+modelsummary(bayesian_model_commercial_brms)
+
+# Extract random effects using brms
+ranef_bayesian_model_commercial_brms <- ranef(bayesian_model_commercial_brms)
+```
+
+the base line probability of having a country that is fishing for commercial purpose is log of 0.838 which is 69.8%.
+which means that there is 69.8% probability that the country is fishing for commercial purposes, without even knowing the country.
+
+```{r}
+# Extract random effects for Country
+random_effects_df <- as.data.frame(ranef_bayesian_model_commercial_brms$Country)
+
+# Check the column names (they are already known: "Estimate.Intercept", etc.)
+names(random_effects_df)
+
+# Now, plot the distribution of the random intercepts for Country
+library(ggplot2)
+
+ggplot(random_effects_df, aes(x = Estimate.Intercept)) + 
+  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black", alpha = 0.7) +
+  labs(title = "Distribution of Random Intercepts for Country",
+       x = "Random Intercept (Country)",
+       y = "Frequency") +
+  theme_minimal()
+
+```
+```{r}
+# Extract random effects for Country
+random_effects_df <- as.data.frame(ranef(bayesian_model_commercial_brms)$Country)
+
+# Add the country names (indexing row names)
+random_effects_df$Country <- rownames(random_effects_df)
+
+# Check the structure to confirm the data is correct
+head(random_effects_df)
+
+# Extract the fixed intercept (the global intercept)
+fixed_intercept <- fixef(bayesian_model_commercial_brms)[1]  # First element corresponds to the intercept
+
+# Add the random intercepts to the fixed intercept to get the log-odds
+random_effects_df$log_odds <- fixed_intercept + random_effects_df$Estimate.Intercept
+
+# Compute predicted probabilities using the log-odds
+random_effects_df$predicted_prob <- 1 / (1 + exp(-random_effects_df$log_odds))
+
+# Plot the predicted probabilities for Commercial catches across countries
+library(ggplot2)
+
+ggplot(random_effects_df, aes(x = Country, y = predicted_prob)) + 
+  geom_bar(stat = "identity", fill = "lightblue", color = "black") +
+  labs(title = "Predicted Probability of Commercial Catch by Country",
+       x = "Country",
+       y = "Predicted Probability") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  # Rotate country names for readability
+  theme_minimal()
+
+```
 
@@
 ## Weaknesses and next steps
+Korea didn't report data until 1969 so they are left out of the first model in general.
 
@@
 
-# Appendix 2
+# Appendix 2 (datasheet)
+
+**1. For what purpose was the dataset created?**
+The dataset was originally created by North Pacific Anadormous Fish Commission [@npafc_statistics] created this dataset to see how much fishing of trout, tuna and salmon was done in international waters.
+They strictly prohibits mass fishing of these highly demanded fishes and hense the dataset.
+We use the dataset to analyse the number of these fishes caught and see if there is a reason for NPAFC to be worried, and if there is a solution to this.
+
+**2.What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)?**
+The instances of these data is either numeric or catagorical.
+Thee first couple of columns tells you the country the data is from, where the fishes were hunted as well as getting to specific regions.
+Why they were hunted and what the unit were, was also in the data set.
+The majority of the dataset is numbers expressing how much fish was hunted, either in thousands or tonnes.
+
+**3.Is any information missing from individual instances?**
+There are several instances of data missing.
+Most are from the number of fishes column as some of the data goes back to 1925 but, not all of them so, there are some rows of data where the total number of fishes collected in 1928, for example, is not present.
+
+**4..
+How was the data associated with each instance acquired?
+Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/ derived from other data (for example, part-of-speech tags, model-based guesses for age or language)?**
+
+This dataset was reported by the subjects of each country. Meaning this organisation asked the government of Canada, US, Korea, Russia and Japan, send in the numbers and they compiled this data.
 

### 2024-11-27 paper/paper.qmd
paper

@@
 subtitle: "Using Baysian Modeling, to find that number of fish caught has increased"
-author: 
-  - Shreya Sakura Noskor
+author: "Shreya Sakura Noskor"
 thanks: "Code and data are available at: [https://github.com/NotSakura/FisheriesData.git](https://github.com/NotSakura/FisheriesData.git)."
@@
 Understanding this trend is important for managing fish populations and ensuring the sustainability of the ocean's resources.
-This is important as the Pacific is full of different kind of fish like, tuna and salmon that are essential in the food chain, but are being overhunted for food and recreational purposes."
+This is important as the Pacific is full of different kind of fish like, trout and salmon that are essential in the food chain, but are being overhunted for food and recreational purposes."
 format:
@@
 data <- read.xlsx('../data/01-raw_data/NPAFC_Catch_Stat-1925-2023.xlsx', sheet = 1) 
-print(data)
 colnames(data) <- data[1, ]
@@
   filter(!is.na(`2022`) & `2022` != 0)
+
+cleaned_pac2 <-cleaned_pac2 %>%
+  select(-Species)
 
@@
 
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+combined_data <- combined_data %>%
+  mutate(across(`1925`:`2023`, as.numeric))
+koreadata <- koreadata %>%
+  mutate(across(`1925`:`2023`, as.numeric))
+
+combined_data <- combined_data  %>%
+  select(-`Whole Country/Province/State`, -`Data Type`, -`Reporting Area`, -Species, -`Catch Type`)
+koreadata <- koreadata  %>%
+  select(-`Whole Country/Province/State`, -`Data Type`, -`Reporting Area`, -Species, -`Catch Type`)
+
+long_data_combined <- combined_data %>%
+  pivot_longer(
+    cols = starts_with("19") | starts_with("20"),  # Select columns starting with "19" or "20"
+    names_to = "Year",
+    values_to = "Catch"
+  ) %>%
+  mutate(Year = as.numeric(Year))
+
+long_data_korea <- koreadata %>%
+  pivot_longer(
+    cols = starts_with("19") | starts_with("20"),  # Select columns starting with "19" or "20"
+    names_to = "Year",
+    values_to = "Catch"
+  ) %>%
+  mutate(Year = as.numeric(Year))
+```
+
 # Introduction
+
+Estimand is the number of fishes that were caught by each country, for every year. Or more specifically what the rate was and if they are more likely to be fished for commercial purposes.
 
@@
 
-China, Japan, Russia
-
 ## Overview
 
+The data was downloaded from @npafc_statistics and was cleaned using R [@citeR]. The data was read using @openexcel and 
+@readxl, while the data was cleaned using @tidy, @tidy, @janitor, @dplyr, @knitr. The data was modeled using @modelsummary, @broom, @rstanarm, and @brms.
+
+To download the data go to [NPFAC's official data portal](https://www.npafc.org/statistics/) and look for "NPAFC Catch Statistics (updated 28 June 2024)".
+Click on that link to get the csv file containing all the data.
+
 ## Methodology and Measurement
+NPAFC has this data to download from their website.
+The way they gathered this data was that they are an inter-government organisation so they have access to government data based on how much fish were hunted in the respective countries.
+The countries in the data include Canada, Russia, Korea, Japan and Unites States of America.
+The way each of these countries measured this data was that when fish are being caught on international waters and report it to each other.
+This is strictly enforced, especially after the fall of salmon and trout population in the Pacific, majorly due to environment purposes.
+
+This paper look at multiple variables.
+First, variable we look at is Country which are either "Canada", "Russia", "Korea", "Japan", and "United States", all representing the countries that are members of this organization.
+This data was left unchanged, Next variables we look at is "Whole Country/Province/State", where the instances of these variables may either be Whole country, or the different states or provinces that was fishing and gathered that data.
+So for example, if the value was British Columbia then the corresponding number of fishes caught reported is the number of fish caught by the province.
+Throughout our data we filtered for the "Whole country" value, assuming that the numbers in each province and/or state would add up to the number in "Whole country" (which it did).
+This was because we were more interested in comparing the fishing trends between countries rather then within a country.
+The next variable is "Reporting Area" which accounts for where the fishes were caught.
+This was also filtered by "Whole country" due to the previous reasoning.
+The next variable that we filtered was "Species" which contained "Cherry", "Chinook", "Chum", "Coho", "Pink", "Sockeye", "Steelhead" and "Total".
+These are all types of salmon except for Steelhead which is a trout and "Total" which represents all the fishes that were hunted.
+Although there is a lot of interesting information to uncover if we did a deeper analysis on each fish, but, we decided that the best way to compare the fishing trends between countries would be to just look at the total fishes caught.
+The next variable we used and actually analyse is the "catch type".
+This tells us whether is the fishes were caught for commercial purposes (caught for profit purposes like selling), sporting purposes (which means they were caught recreationaly) or subsistence purposes (which means they were caught to provide food, not as a profit).
+The last variable we filtered was "Data Type" which was the unit that these numbers were reported in.
+There was Numbers in 1000s or Round weight in metric tonne.
+We chose to filter with numbers in 1000s as the other option was done only by the US, who provided both units.
+
+Thoese were the variables that we filtered or focused on but there is one more variable that is important for this paper; the number of fishes caught.
+That number in the raw data was provided as the year as a column and the corresponding value as the number of fishes caught.
+This format meant that when we are creating models or graphs it is very difficult to work with it.
+And so we actually shifted, rather pivotted the table so that analysis is easier.
+To pivot the table we created 2 new rows to the dataset, "Year" and "Catch".
+The year corresponds to the column's title which is the year this data is for and the "catch" refers to the number of fishes that were caught in that year, in that country.
+This helped significantly with making the models and such.
 
@@
 
-we make the assumption that the coulms that say whole country it also include the provinces and different areas number as well.
+we make the assumption that the coloumns that say whole country it also include the provinces and different areas number as well.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-long_comb
+
+ggplot(long_data_combined, aes(x = Year, y = Catch, color = Country)) +
+  geom_line(size = 1, alpha = 0.7) +  # Line plot with some transparency
+  geom_point(size = 2, alpha = 0.8) +  # Points with smaller size
+  labs(title = "Catch over Time by Country", x = "Year", y = "Catch") +
+  scale_y_continuous(labels = scales::comma) +  # Format y-axis with commas
+  theme_minimal() +
+  theme(legend.title = element_blank())
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-korea
+
+ggplot(long_data_korea, aes(x = Year, y = Catch, color = Country)) +
+  geom_line(size = 1, alpha = 0.7) +  # Line plot with some transparency
+  geom_point(size = 2, alpha = 0.8) +  # Points with smaller size
+  labs(title = "Catch over Time by Country", x = "Year", y = "Catch") +
+  scale_y_continuous(labels = scales::comma) +  # Format y-axis with commas
+  theme_minimal() +
+  theme(legend.title = element_blank())
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-total_bar
+
+# Load necessary libraries
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+
+# Create a complete grid of all possible combinations of `Country` and `Catch Type`
+complete_data <- expand.grid(
+  `Catch Type` = unique(cleaned_pac2$`Catch Type`),
+  Country = unique(cleaned_pac2$Country)
+)
+
+aggregated_data <- cleaned_pac2 %>%
+  group_by(`Catch Type`, Country) %>%
+  summarise(
+    total_catch = sum(as.numeric(`2022`), na.rm = TRUE), # Convert to numeric
+    .groups = "drop"
+  )
+
+# Replace "United States" with "US" in the Country variable
+full_data <- complete_data %>%
+  left_join(aggregated_data, by = c("Catch Type", "Country")) %>%
+  mutate(total_catch = ifelse(is.na(total_catch), 0, total_catch)) %>%  # Replace NA with 0 for missing catch data
+  mutate(Country = ifelse(Country == "United States", "US", Country))  # Abbreviate United States to US
+
+# Create the plot with the aggregated data and numbers on the bars
+ggplot(full_data, aes(x = `Catch Type`, y = total_catch, fill = Country)) +
+  geom_bar(stat = "identity", position = "dodge") +  # Bar plot with grouped bars by country
+  geom_text(aes(label = paste(Country, ":\n", round(total_catch, 0))),  # Newline between Country and catch number
+            position = position_dodge(width = 0.9), 
+            vjust = -0.1,  # Adjust vertical position of text
+            size = 2) +  # Smaller font size for the numbers
+  labs(title = "Catch Data by Country and Catch Type (2022)", 
+       x = "Catch Type", 
+       y = "Catch Data") +
+  scale_y_continuous(labels = scales::comma) +
+  theme_minimal() +
+  theme(legend.title = element_blank())
+
+```
 
@@
 # Model
+Here we model the data in 2 ways.
+The first way is to model the rate at which the fishes are being caught for any country.
+And the second model looks at what is the probability that a country is fishing for commercial purposes.
 
@@
 #| message: false
-temp <- cleaned_pac  %>%
-  select(-`Whole Country/Province/State`, -`Data Type`, -`Reporting Area`, -Species, -`Catch Type`)
-temp
+library(tidyr)
+library(rstanarm)
 
+#model_formula <- Catch ~ Year + (1 | Country)
+
+#bayesian_model <- stan_glmer(
+#  formula = model_formula,
+#  data = long_data_combined,
+#  family = gaussian(),  # Adjust this if using a different distribution for your data
+#  prior = normal(0, 2.5, autoscale = TRUE),
+#  prior_intercept = normal(0, 2.5, autoscale = TRUE),
+#  seed = 123,
+#  adapt_delta = 0.95,
+ # cores = 4
+#)
+#saveRDS(
+#  bayesian_model,
+#  file = "../models/bayesian_model1.rds"
+#)
+bayesian_model <-
+  readRDS(file = here::here("models/bayesian_model1.rds"))
 ```
+
+Model 2
 
@@
 #| message: false
-library(tidyr)
-library(dplyr)
-
-# Convert columns that should be years to numeric
-temp <- temp %>%
-  mutate(across(`1925`:`2023`, as.numeric))
-
-long_data <- temp %>%
-  pivot_longer(
-    cols = starts_with("19") | starts_with("20"),  # Select columns starting with "19" or "20"
-    names_to = "Year",
-    values_to = "Catch"
-  ) %>%
-  mutate(Year = as.numeric(Year))
-
-library(rstanarm)
-
-model_formula <- Catch ~ Year + (1 | Country)
-
-bayesian_model <- stan_glmer(
-  formula = model_formula,
-  data = long_data,
-  family = gaussian(),  # Adjust this if using a different distribution for your data
-  prior = normal(0, 2.5, autoscale = TRUE),
-  prior_intercept = normal(0, 2.5, autoscale = TRUE),
-  seed = 123,
-  adapt_delta = 0.95,
-  cores = 4
-)
-
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-modelsummary(bayesian_model)
-```
-
-Model 2
-
-```{r}
 
@@
 # Specify priors
-priors <- c(
-  set_prior("normal(0, 2.5)", class = "Intercept"),   # Prior for the intercept
-  set_prior("normal(0, 2.5)", class = "sd")           # Prior for the standard deviation of random effects (Country)
-)
+#priors <- c(
+#  set_prior("normal(0, 2.5)", class = "Intercept"),   # Prior for the intercept
+#  set_prior("normal(0, 2.5)", class = "sd")           # Prior for the standard deviation of random effects (Country)
+#)
+
+#cleaned_pac2$Is_Commercial <- ifelse(cleaned_pac2$`Catch Type` == "Commercial", 1, 0)
 
 # Fit the model using brms
-bayesian_model_commercial_brms <- brm(
-  formula = Is_Commercial ~ (1 | Country),  # Random effect for Country
-  data = cleaned_pac2,
-  family = bernoulli(),                    # Bernoulli family for binary outcomes
-  prior = priors,                          # Use the prior argument here
-  seed = 123
-)
+#bayesian_model_commercial_brms <- brm(
+#  formula = Is_Commercial ~ (1 | Country),  # Random effect for Country
+#  data = cleaned_pac2,
+#  family = bernoulli(),                    # Bernoulli family for binary outcomes
+#  prior = priors,                          # Use the prior argument here
+ # seed = 123
+#)
 
-# Check the model summary
-modelsummary(bayesian_model_commercial_brms)
+#saveRDS(
+#  bayesian_model_commercial_brms,
+#  file = "../models/bayesian_model2.rds"
+#)
 
+bayesian_model_commercial_brms <-
+  readRDS(file = here::here("models/bayesian_model2.rds"))
 # Extract random effects using brms
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
 # Extract random effects for Country
@@
 ```
+
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
 # Extract random effects for Country
@@
 
+# Results
+
 ## Model Results
 
-# Results
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig_model1
+#| tbl-cap: "baysian model summary for predicting the rate of fishes being caught"
+
+modelsummary::modelsummary(
+  list(
+    "First model" = bayesian_model
+  ),
+  statistic = "mad",
+  fmt = 2
+)
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig_model2
+#| tbl-cap: "baysian model summary for predicting the probability of country fishing for commercial purposes."
+modelsummary::modelsummary(
+  list(
+    "Second model" = bayesian_model_commercial_brms
+  ),
+  statistic = "mad",
+  fmt = 2
+)
+```
 
@@
 
-**1. For what purpose was the dataset created?**
-The dataset was originally created by North Pacific Anadormous Fish Commission [@npafc_statistics] created this dataset to see how much fishing of trout, tuna and salmon was done in international waters.
+1. For what purpose was the dataset created?
+The dataset was originally created by North Pacific Anadormous Fish Commission [@npafc_statistics] created this dataset to see how much fishing of trout and salmon was done in international waters.
 They strictly prohibits mass fishing of these highly demanded fishes and hense the dataset.
@@
 
-**2.What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)?**
+2.What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)?
 The instances of these data is either numeric or catagorical.
@@
 
-**3.Is any information missing from individual instances?**
+3.Is any information missing from individual instances?
 There are several instances of data missing.
@@
 
-**4..
+4.
 How was the data associated with each instance acquired?
-Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/ derived from other data (for example, part-of-speech tags, model-based guesses for age or language)?**
+Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/ derived from other data (for example, part-of-speech tags, model-based guesses for age or language)?
 

### 2024-11-27 scripts/06-model_data.R

@@ -76,3 +76,6 @@ saveRDS(
 )
 
 
+
+
+

### 2024-11-27 paper/paper.qmd
fixed the latex issue where model summary wont show up "Just installed the tinytax stuff on qerto page"

@@
     geometry: margin = 1in
+header-includes:
+  - \usepackage{float} 
+  - \floatplacement{table}{H}
 number-sections: true
+include-in-header: 
+  text:
+    \renewcommand{\abstractname}{Abstract}
+bibliography: references.bib
 toc: true
-bibliography: references.bib
 ---
@@
 
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`. We also use `brm` package from @brms
+
 ## Model Step- Up
@@
 #| message: false
-#| label: fig_model1
+#| label: tbl-model1
 #| tbl-cap: "baysian model summary for predicting the rate of fishes being caught"
@@
 #| message: false
-#| label: fig_model2
+#| label: tbl-model2
 #| tbl-cap: "baysian model summary for predicting the probability of country fishing for commercial purposes."
+
 modelsummary::modelsummary(

### 2024-11-29 paper/paper.qmd
model section and data section done. Still need to do appendix, discussion, results, and intro. update sketches and LLM. And try to figure out what a model card is.

@@
 
-This paper look at multiple variables.
-First, variable we look at is Country which are either "Canada", "Russia", "Korea", "Japan", and "United States", all representing the countries that are members of this organization.
-This data was left unchanged, Next variables we look at is "Whole Country/Province/State", where the instances of these variables may either be Whole country, or the different states or provinces that was fishing and gathered that data.
+This paper look at multiple variables. We will go through them one by one: 
+- First, variable we look at is Country which are either "Canada", "Russia", "Korea", "Japan", and "United States", all representing the countries that are members of this organization.
+This data was left unchanged.
+- Next variables we look at is "Whole Country/Province/State", where the instances of these variables may either be Whole country, or the different states or provinces that was fishing and gathered that data.
 So for example, if the value was British Columbia then the corresponding number of fishes caught reported is the number of fish caught by the province.
@@
 This was because we were more interested in comparing the fishing trends between countries rather then within a country.
-The next variable is "Reporting Area" which accounts for where the fishes were caught.
+- The next variable is "Reporting Area" which accounts for where the fishes were caught.
 This was also filtered by "Whole country" due to the previous reasoning.
@@
 Although there is a lot of interesting information to uncover if we did a deeper analysis on each fish, but, we decided that the best way to compare the fishing trends between countries would be to just look at the total fishes caught.
-The next variable we used and actually analyse is the "catch type".
+- The next variable we used and actually analyse is the "catch type".
 This tells us whether is the fishes were caught for commercial purposes (caught for profit purposes like selling), sporting purposes (which means they were caught recreationaly) or subsistence purposes (which means they were caught to provide food, not as a profit).
-The last variable we filtered was "Data Type" which was the unit that these numbers were reported in.
+- The last variable we filtered was "Data Type" which was the unit that these numbers were reported in.
 There was Numbers in 1000s or Round weight in metric tonne.
@@
 
-Thoese were the variables that we filtered or focused on but there is one more variable that is important for this paper; the number of fishes caught.
+These were the variables that we filtered but the variable we actually analyse is the number of fishes caught.
 That number in the raw data was provided as the year as a column and the corresponding value as the number of fishes caught.
@@
 
-we make the assumption that the coloumns that say whole country it also include the provinces and different areas number as well.
+we make the assumption that the columns that say whole country it also include the provinces and different areas number as well.
 
-```{r}
+```{r, fig.height= 3.5, fig.width= 7.5}
 #| echo: false
@@
 #| label: fig-long_comb
+#| fig-cap: "Catch over Time by Country"
 
@@
   geom_point(size = 2, alpha = 0.8) +  # Points with smaller size
-  labs(title = "Catch over Time by Country", x = "Year", y = "Catch") +
+  labs(x = "Year", y = "Catch") +
+  scale_y_continuous(labels = scales::comma) +  # Format y-axis with commas
+  theme_minimal() +
+  theme(legend.title = element_blank())
+
+```
+From @fig-long_comb we are able to see the number of fishes caught in each country as time goes by.
+They fluctuate a lot because over fishing in one year means that the next year there is not a lot of fish left as not as many survived to repopulate.
+However we see that there is a increase in the number of fishes caught for the US and Russia.
+Meanwhile, Japan and Canada are declining.
+This is interesting as Japan is surrounded by the Pacific Ocean more than all of these other countries and Canada coming in a close second.
+An educated guess as to why this may be the case is that, the US and Russia may be fishing more as they have a stronger fishing industry with the technology and money to fund longer trips in the ocean.
+
+```{r, fig.height= 3.5, fig.width= 6.5}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-korea
+#| fig-cap: "Catch over Time by Country"
+
+ggplot(long_data_korea, aes(x = Year, y = Catch, color = Country)) +
+  geom_line(size = 1, alpha = 0.7) +  # Line plot with some transparency
+  geom_point(size = 2, alpha = 0.8) +  # Points with smaller size
+  labs( x = "Year", y = "Catch") +
   scale_y_continuous(labels = scales::comma) +  # Format y-axis with commas
@@
 
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-korea
+In @fig-long_comb we can barely see the fluctuation for Korea; it seems almost constant.
+Hense we graph it sepreatly here in @fig-korea.
+We see that like the other countries in our data, they also fluctuate in numbers.
+From this an guess may be that they are decreasing the number of fishes they catch but, our guess is quite the opposite.
+We think that in 2023 it was just low due to one of the fluctuations where they fished too much the previous year.
+We expect a rise in the next 2 or 3 years.
 
-ggplot(long_data_korea, aes(x = Year, y = Catch, color = Country)) +
-  geom_line(size = 1, alpha = 0.7) +  # Line plot with some transparency
-  geom_point(size = 2, alpha = 0.8) +  # Points with smaller size
-  labs(title = "Catch over Time by Country", x = "Year", y = "Catch") +
-  scale_y_continuous(labels = scales::comma) +  # Format y-axis with commas
-  theme_minimal() +
-  theme(legend.title = element_blank())
-
-```
-
-```{r}
+```{r, fig.height= 3.5, fig.width= 7.5}
 #| echo: false
@@
 #| label: fig-total_bar
+#| fig-cap: "Catch Data by Country and Catch Type (2022)"
 
@@
             size = 2) +  # Smaller font size for the numbers
-  labs(title = "Catch Data by Country and Catch Type (2022)", 
+  labs( 
        x = "Catch Type", 
@@
 ```
+@fig-total_bar is a great way to analyse the different reasons why these fishes are caught.
+We narrowed the year to 2022 as we predict the fishing industry has recovered after covid and will show accurate results.
+Here we see that most fishes are caught due to commercial purposes where they sell the fish rather then for sporting or subsistence purposes.
+It also seems that Russia fishes the most out of the Norther Pacific region with its number being almost doubled from the second most country to fish, the US.
+Russia has fished around 309,142,000 fishes in the year 2022 next to US who has fished around 166,114,000.
+With these numbers, no wonder we see fluctuation in the number of fishes caught in @fig-long_comb.
 
-## Outcome and Predictor variables
-
+\newpage 
 # Model
@@
 
-## Model Step- Up
+## Number of Fishes Caught (dependent on Year and Country)
+
+Define $Catch_{ij}$ to be the number of fishes caught.
+We are trying to see how will $Catch_{ij}$ change as we increase $Year$ which is our other variable.
+We are finding the correlation with the countries in mind which is why we account for the random effect of the country.
+
+\begin{align} 
+\mbox{Catch}_{ij} &\sim \mbox{Normal}(\mu_{ij}, \sigma^2) \\
+\mu_{ij} &= \beta_0 + \beta_1 \times \mbox{Year}_{ij} + u_j \\
+u_j &\sim \mbox{Normal}(0, \sigma_{\mbox{Country}}^2) \\
+\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
+\end{align}
+
+### Model Justification
+- $Catch_{ij}$ is the number of fishes caught, the variable we are modeling
+- $Year$ is the independent variable that shows the year
+- $u_j$ is the random effect of the countries. This is important as the number of fishes caught differentiate between country so to get a weighted result we add this variable. 
+- $\beta_0$ and $\beta_1$ tells us the intercept and the slope respectively, of the Bayesian Generalised Linear Mixed Model.
+
+We predict that even thought 2 (may be 3) out of the 5 countries in the data set have a positive trend in terms of the number of fish being caught (@fig-long_comb), because the rate at which Russia and USA is increasing is way higher than Japan, Korea and Canada.
+This can be seen visually.
+So I think that the rate of increase is going to be positive.
 
@@
 
-Model 2
+## To be Commercial or Not to be Commercial
+Define $IsCommercial_{ij}$ to be the probability that the fishes caught were commercial (for buisness purposes).
+We are trying to see how will $IsCommercial_{ij}$ change as we change $Country$.
+We esstially use, again, a Baysian Generalised Linear Mixture Model to see how does the likelihood of fishes being caught for commercial purposes varies from country to country.
+
+\begin{align} 
+\mbox{IsCommercial}_{ij} &\sim \mbox{Bernoulli}(p_{ij}) \\
+\mbox{logit}(p_{ij}) &= \beta_0 + u_j \\ 
+\mbox{u}_j &\sim \mbox{Normal}(0, \sigma_{\mbox{Country}}^2) \\
+\beta_0 &\sim \mbox{Normal}(0, 2.5) \\ 
+\sigma_{\mbox{Country}} &\sim \mbox{Normal}(0, 2.5)
+\end{align}
+
+### Model Justification
+- $IsCommercial_{ij}$ is the probability that the fishes caught were commercial. This is what we are modelling using Bernoulli Distribution. 
+- $Country$ is the independent variable that shows the countries. 
+- $u_j$ is the random effect of the countries. This is important as the number of fishes caught commercially is different between countries. 
+- $\beta_0$  tells us the intercept so, what is the likelihood of it being commercial for a "baseline" country. This is essentially the average of all the intercepts for all the countries.  
+Note that this model doesn't look at the number of fishes caught rather the number of occurrences in the year 2022 where fishes were caught commercially.
+We focus on the year 2022 because it is recent data and modeling with all years between 1925 to 2023 would have not given us a result that was relevant to us.
+We would also loose information as the liklihood may have not been as high in the 1900s due to the lack of equipment.
+
+We predict that the outcome or rather the probability that the fishes caught were commercially would be positive.
+This is because as we were looking at the graphs in the data section, we notice that @fig-total_bar shows most of the fishes caught were commercially and then as a sport and then as a subsistence.
 
@@
 
+### Model justification
+
 the base line probability of having a country that is fishing for commercial purpose is log of 0.838 which is 69.8%.
 which means that there is 69.8% probability that the country is fishing for commercial purposes, without even knowing the country.
+
+# Results
+
+## Model Results
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| label: tbl-modelSum
+#| layout-ncol: 2
+#| tbl-cap: "Model Summaries"
+#| tbl-subcap: ["baysian model summary for predicting the probability of country fishing for commercial purposes.", "baysian model summary for predicting the rate of fishes being caught"]
+
+modelsummary::modelsummary(
+  list(
+    "Second model" = bayesian_model_commercial_brms
+  ),
+  statistic = "mad",
+  fmt = 2
+)
+
+modelsummary::modelsummary(
+  list(
+    "First model" = bayesian_model
+  ),
+  statistic = "mad",
+  fmt = 2
+)
+```
+
+## Other statistics
 
@@
 #| message: false
+#| results: hide
+#| label: fig-randomInter
+#| fig-cap: "Distribution of Random Intercepts for Country"
 # Extract random effects for Country
@@
 ggplot(random_effects_df, aes(x = Estimate.Intercept)) + 
-  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black", alpha = 0.7) +
-  labs(title = "Distribution of Random Intercepts for Country",
+  geom_histogram(binwidth = 0.1, fill = "grey", color = "black", alpha = 0.7) +
+  labs(
        x = "Random Intercept (Country)",
@@
 #| message: false
+#| results: hide
+#| label: fig-predProb
+#| fig-cap: "Predicted Probability of Commercial Catch by Country"
 # Extract random effects for Country
@@
 ggplot(random_effects_df, aes(x = Country, y = predicted_prob)) + 
-  geom_bar(stat = "identity", fill = "lightblue", color = "black") +
-  labs(title = "Predicted Probability of Commercial Catch by Country",
+  geom_bar(stat = "identity", fill = "grey", color = "black") +
+  labs(
        x = "Country",
@@
 
-```
-
-### Model justification
-
-# Results
-
-## Model Results
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: tbl-model1
-#| tbl-cap: "baysian model summary for predicting the rate of fishes being caught"
-
-modelsummary::modelsummary(
-  list(
-    "First model" = bayesian_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: tbl-model2
-#| tbl-cap: "baysian model summary for predicting the probability of country fishing for commercial purposes."
-
-modelsummary::modelsummary(
-  list(
-    "Second model" = bayesian_model_commercial_brms
-  ),
-  statistic = "mad",
-  fmt = 2
-)
 ```

### 2024-11-30 paper/paper.qmd
model card added, appendix A finished so all appendi done (unles i add to appendix B). Finished results. Left: Discussion, Intro, Sketches and LLM

@@
 title: "Analysing the Fish Hunting Numbers in North Pacific Ocean"
-subtitle: "Using Baysian Modeling, to find that number of fish caught has increased"
+subtitle: "Using Bayesian Modeling, to find that number of fish caught has increased"
 author: "Shreya Sakura Noskor"
@@
 
-## Number of Fishes Caught (dependent on Year and Country)
+## Number of Fishes Caught (dependent on Year and Country) {#sec-mod1}
 
@@
 
-## To be Commercial or Not to be Commercial
+## To be Commercial or Not to be Commercial {#sec-mod2}
 Define $IsCommercial_{ij}$ to be the probability that the fishes caught were commercial (for buisness purposes).
 We are trying to see how will $IsCommercial_{ij}$ change as we change $Country$.
-We esstially use, again, a Baysian Generalised Linear Mixture Model to see how does the likelihood of fishes being caught for commercial purposes varies from country to country.
+We esstially use, again, a Bayesian Generalised Linear Mixture Model to see how does the likelihood of fishes being caught for commercial purposes varies from country to country.
 
@@
 
-### Model justification
-
-the base line probability of having a country that is fishing for commercial purpose is log of 0.838 which is 69.8%.
-which means that there is 69.8% probability that the country is fishing for commercial purposes, without even knowing the country.
-
 # Results
@@
 #| tbl-cap: "Model Summaries"
-#| tbl-subcap: ["baysian model summary for predicting the probability of country fishing for commercial purposes.", "baysian model summary for predicting the rate of fishes being caught"]
+#| tbl-subcap: ["bayesian model summary for predicting the probability of country fishing for commercial purposes.", "bayesian model summary for predicting the rate of fishes being caught"]
 
@@
 
+In @tbl-modelSum-1 we see the model summary for Model 1 which was finding the probability of a country fishing for commercial purposes.
+And in @tbl-modelSum-2 we see the results for model 2 where we look at the rate at which fish are being hunted for the "baseline" country.
+There are a lot of numbers here that we don't need to look at to see if our predictions were right.
+We only need to know the intercepts and the slopes or as shown in the model $\beta_0$ and $\beta_1$.
+In @tbl-modelSum-1 we see that the log likelihood of a "baseline" country fishing for commercial purpose is 0.84 which is 69.8%.
+This means that there is 69.8% probability that the country is fishing for commercial purposes, without even knowing the country.
+The "sd country Intercept" just tells you how much variation there is in that result.
+Next in @tbl-modelSum-2 we see that the intercept is -1,367,335 which tells you that at Year 0 somehow a baseline country fished negative -1,367,335 fish.
+We will discuss more about this number and interpret it in the next section (@sec-disc).
+The "Year = 721.74" tells you that with every increase in year there is about 721,000 more fish that are being caught.
+And finally similar to the other model summary "sd country $\times$ Intercept Intercept" tells you how much variation there is in the results.
+We don't need these variation values unless we are analysing the model (and not the data).
+
 ## Other statistics
@@
 #| results: hide
-#| label: fig-randomInter
-#| fig-cap: "Distribution of Random Intercepts for Country"
+##| label: fig-randomInter
+##| fig-cap: "Distribution of Random Intercepts for Country"
 # Extract random effects for Country
@@
 # Now, plot the distribution of the random intercepts for Country
-library(ggplot2)
+#library(ggplot2)
 
-ggplot(random_effects_df, aes(x = Estimate.Intercept)) + 
-  geom_histogram(binwidth = 0.1, fill = "grey", color = "black", alpha = 0.7) +
-  labs(
-       x = "Random Intercept (Country)",
-       y = "Frequency") +
-  theme_minimal()
+#ggplot(random_effects_df, aes(x = Estimate.Intercept)) + 
+#  geom_histogram(binwidth = 0.1, fill = "grey", color = "black", alpha = 0.7) +
+#  labs(
+#       x = "Random Intercept (Country)",
+#       y = "Frequency") +
+#  theme_minimal()
 
@@
   geom_bar(stat = "identity", fill = "grey", color = "black") +
+  geom_text(aes(label = round(predicted_prob, 2)), vjust = -0.5) +  # Add text labels rounded to 2 decimals
   labs(
@@
 
-# Discussion
+We also calculate the predicted probability of a country fishing commercially for all 5 countries in the NPFAC.
+In @fig-predProb we see that Canada, US and Russia has a 40% probability that they fish commercially, Japan and Korea both have 97% probability.
+We will see why that is the case in the next section (@sec-disc).
+
+\newpage
+
+# Discussion {#sec-disc}
+
+## Story of the Data
 
@@
 
-# Appendix 1
+# Fisheries and Surveys/Observational Data {#sec-A}
 
-# Appendix 2 (datasheet)
+The data that we gathered for this paper is definitely observational data.
+This is because the in the NPFAC's official website they state that they gather the data by contacting each member of government and asking them to submit this data.
+In our own Canadian government, we see that we have Fish Monitoring Policy where people are expected to report their catch data if they caught a lot [@govFish].
+For example, if some one has fished a rare specie of fish or a large number of fishes were caught they must report it to the government.
+This is how the government gains the data and passes it on to NPFAC.
 
-1. For what purpose was the dataset created?
-The dataset was originally created by North Pacific Anadormous Fish Commission [@npafc_statistics] created this dataset to see how much fishing of trout and salmon was done in international waters.
-They strictly prohibits mass fishing of these highly demanded fishes and hense the dataset.
+The common concern with datasets of our type is often 2 paradoxes know as Simpson's Paradox and Berkson's Paradox.
+We assume that Simpson's paradox does come into play for our model in @sec-mod1.
+Simpson's Paradox is when we have relationship between subsets of data show a result but when we model the entire result together, the result differs.
+This can be seen in our first model because we see that in @fig-long_comb that different countries have a different rate of change for the number of catches with respect to year.
+We see that Japan and Canada has a decreasing rate.
+However the over all model outputs a positive rate of change.
+This shows that the subset of the data has a different result then when we combine the entire data.
+So this paradox does apply to us.
+Next we also say that the Berkson's Paradox also affects us as in our data set we only take a look at the data for Salmon and Trout Population.
+However, what the dataset doesn't account is the other fish species that are also fished.
+Yes it is true that over fishing is an issue but the result that our model shows may not be the most accurate number due to the fact that this key information is left out.
+The @popFish shows that salmon, tuna and trout are one of the most fished anadromous.
+And @npafc_species shows the other anadromous fishes that are in the North Pacific Ocean.
+
+We observed that one of the reasons why these paradox applies to us is because our data doesn't sample the entire data.
+The way they gather their data is similar to an Integrated method (mentioned in Stantcheva's supplemental appendix in @sampling).
+Although they are not an open community where anyone who is a member of the community may be rewarded for their participation in giving their data, all of the countries is under an organisation and they give their data as they all actively participate in the organisations goal for sustainable fishing.
+The issue we were talking about earlier- the one that causes these paradoxes to affect our data - is the fact that the way this organisation collects data also presents a sampling bias.
+A sampling bias is when the data collected is not a representitive of the entire population.
+This can be caused when data is collected only from one place and we ignore all other factors.
+This is what happens here.
+Because NPFAC only consists of the 5 countries in the data (Japan, Korea, Russia, US, Canada), our data has gaps for the number of fishes caught by other countries that fish in the North Pacific Ocean such as other East Asian countries like China, Phillipines and Taiwan.
+Realistically China has a large fishing industry as well so our analysis would look very different when we include their data.
+Another thing we are missing is the fact that there are other Anadromous fishes in the Pacific that could also skew the data and causes the Berkson's Paradox.
+
+We will now move on from critiquing the data to our analysis.
+Our modeling is a prime example of convenience sampling.
+This is when we filter data to our own rules and analyse those results because it is the most convenient for us.
+In this paper it was done for @fig-long_comb.
+This was due to the fact that not a lot of the data dated back to the 1925 so, initially Korea was left out from our data.
+However, because we wanted a chance to compare all 5 countries of the organisation, we combined data from Korea which started 1969.
+We purposefully chose filters that will shorten our data while getting getting the attributes we want to study.
+The effect of this is that our model where we check the rate of change of fishes caught [@sec-mod1], is inaccurate as we don't have the data from 1925 to 1969 for Korea and we filtered the data so much to fit Korea that we loose information like the type of fish and why they were fished (commercial, sporty, subsistence).
+
+Now we will move to the survey analysis.
+Because this data was not gathered through a survey we will simply talk about how we would gather this data.
+It would not be very complicated.
+Our hope is that 6-months before we send the survey we ask the fisher to keep track of their weekly fish caught.
+That they would record the number of fish caught of each type, the area they were in and when they caught them (time of day would be appreciated but the date is minimum).
+Then 6 months after the notice I send out a survey that asks for their information.
+It would be nothing private like the area they live in and how long they spent fishing.
+I would send out this email to corporations for their data as well as regular fisherman that fish for subsistence.
+This would be to make sure there is not sampling bias when collecting the data.
+This was an issue in our original data as we noticed there were a lot more commercial fishes caught and we did not know if this was due to not a lot of fish caught for subsistence or they were not reached out to for their data.
+This would make sure that we have as minimal gaps in our data as possible.
+
+# Data Sheet
+
+*1. For what purpose was the dataset created?*
+The data set was originally created by North Pacific Anadromous Fish Commission [@npafc_statistics] created this dataset to see how much fishing of trout and salmon was done in international waters.
+They strictly prohibits mass fishing of these highly demanded fishes and hence the dataset.
 We use the dataset to analyse the number of these fishes caught and see if there is a reason for NPAFC to be worried, and if there is a solution to this.
 
-2.What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)?
-The instances of these data is either numeric or catagorical.
+*2. What do the instances that comprise the data set represent (for example, documents, photos, people, countries)?*
+The instances of these data is either numeric or categorical.
 Thee first couple of columns tells you the country the data is from, where the fishes were hunted as well as getting to specific regions.
@@
 
-3.Is any information missing from individual instances?
+*3. Is any information missing from individual instances?*
 There are several instances of data missing.
@@
 
-4.
+*4.
 How was the data associated with each instance acquired?
-Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/ derived from other data (for example, part-of-speech tags, model-based guesses for age or language)?
+Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/ derived from other data (for example, part-of-speech tags, model-based guesses for age or language)?*
 
@@
 
+\newpage
+
+# Model Card - North Pacific Anadromous Fish Data
+**Model Details**
+- Model was created by Student at University of Toronto on November 27th 2024. 
+- Bayesian Generalized Linear Model (GLMM)
+- Ver 4 as the other versions tried to take all variables like region and country into consideration but the model was too complex to implement and even more to analyse.
+
+**Intended Use**
+- Analyse the number of Anadromous fish caught in the North Pacific Ocean by looking at if they are commercially caught and the rate at which they were caught. 
+- Intended to see if sustainability measures put on fish hunting has worked
+- Not intended to see what measures can be put into change the outcome of the results seen here (such as laws to prevent the rate being so high)
+
+**Factors**
+- Relevant factors are year, number of fishes caught and if the fishes were caught for commercial purpose or not (true or false value).
+Year is a predetermined value and the countries that reported the data to NPFAC was in charge of identifying the number of fishes they caught and why they caught it.
+Details available in data section of the paper.
+- Evaluation factors or the values that are being reported are rate of number of fish caught [@fig-modelcard-2] and probability that the fish caught was commercial [@fig-modelcard-3].
+These are being evaluated to understand if certain sustainability measures are helping the fish being over hunted.
+
+**Metrics**
+- Measured in Thousands of Fish as given in @npafc_statistics.
+
+**Training Data**
+- Data from @npafc_statistics was used to make/train this model.
+
+**Evaluation Data**
+- We applied the model on the same data as the training set [@fig-modelcard-2] and we see that it is linear due to the fact that we applied the model on it self with the same rate.
+
+**Ethical Considerations**
+- We are only using the observational data.
+Hence we are not actually changing the data themselves.
+The model does not take into account any personal data as the only data provided was the year, the number of fish caught and where they were caught.
+
+**Caveats and Recommendation**
+- Use a data that has no sampling bias. This is because of the reasons discussed in @sec-A. To summaries it, it is because of the Simpson's and Berkson's Paradox that seems to arise.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| label: fig-modelcard
+#| layout-ncol: 3
+#| fig-cap: "Quantitative Analysis"
+#| fig-subcap: ["Bar plot with error bars for predictions and credible intervals by Country", "Plot predictions over time (by Year) for each Country", "Predicted Probability of Commercial Catch by Country"]
+#| results: hide
+
+# Load necessary libraries
+library(rstanarm)
+library(ggplot2)
+
+# Assuming the model is already fitted with stan_glmer like this:
+# model_formula <- Catch ~ Year + (1 | Country)
+# bayesian_model <- stan_glmer(formula = model_formula, data = long_data_combined, family = gaussian(), 
+#                              prior = normal(0, 2.5, autoscale = TRUE), 
+#                              prior_intercept = normal(0, 2.5, autoscale = TRUE), 
+#                              seed = 123, adapt_delta = 0.95, cores = 4)
+
+# Extract posterior predictions from the fitted model
+posterior_preds <- posterior_predict(bayesian_model)
+
+# Summarize the predictions (mean and credible intervals)
+mean_preds <- apply(posterior_preds, 2, mean)  # Mean of posterior predictions for each data point
+lower_ci <- apply(posterior_preds, 2, function(x) quantile(x, probs = 0.025))  # 95% lower credible interval
+upper_ci <- apply(posterior_preds, 2, function(x) quantile(x, probs = 0.975))  # 95% upper credible interval
+
+# Check the number of rows in the original data and posterior predictions
+print(nrow(long_data_combined))  # Should be 495 (or the correct number of rows)
+print(length(mean_preds))  # Should match the number of rows in long_data_combined
+
+# Handle potential mismatch in number of rows (e.g., due to missing values)
+# If there are any missing values, we can subset the data or remove rows with NAs
+long_data_combined_clean <- long_data_combined[complete.cases(long_data_combined$Catch), ]
+
+# Ensure the posterior predictions align with the cleaned data
+# If long_data_combined_clean has a different number of rows than the posterior predictions,
+# subset the predictions accordingly.
+if (nrow(long_data_combined_clean) != length(mean_preds)) {
+  stop("The number of posterior predictions does not match the cleaned data.")
+}
+
+# Add the predictions and credible intervals to the cleaned data
+long_data_combined_clean$mean_pred <- mean_preds[1:nrow(long_data_combined_clean)]
+long_data_combined_clean$lower_ci <- lower_ci[1:nrow(long_data_combined_clean)]
+long_data_combined_clean$upper_ci <- upper_ci[1:nrow(long_data_combined_clean)]
+
+# Visualizing the predictions using ggplot2
+# Bar plot with error bars for predictions and credible intervals by Country
+ggplot(long_data_combined_clean, aes(x = Country, y = mean_pred, fill = Country)) +
+  geom_bar(stat = "identity", position = "dodge", color = "black") +
+  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
+  labs(title = "Model Predictions by Country with 95% Credible Intervals", 
+       x = "Country", 
+       y = "Mean Prediction") +
+  theme_minimal()
+
+# If you want to plot predictions over time (by Year) for each Country
+ggplot(long_data_combined_clean, aes(x = Year, y = mean_pred, color = Country)) +
+  geom_point() +
+  geom_line() +
+  facet_wrap(~ Country, scales = "free_y") +  # Facet by Country
+  labs(title = "Predictions of Catch by Year and Country", x = "Year", y = "Mean Prediction") +
+  theme_minimal()
+
+# Optionally, compute and visualize R-squared or other performance metrics
+# Compute R-squared
+observed_values <- long_data_combined_clean$Catch
+mse <- mean((observed_values - long_data_combined_clean$mean_pred)^2)  # Mean squared error
+r_squared <- 1 - sum((observed_values - long_data_combined_clean$mean_pred)^2) / sum((observed_values - mean(observed_values))^2)
+
+# Print MSE and R-squared
+print(paste("MSE:", mse))
+print(paste("R-squared:", r_squared))
+
+# Optionally, you can add these metrics to the data for further visualization if needed
+long_data_combined_clean$mse <- mse
+long_data_combined_clean$r_squared <- r_squared
+
+ggplot(random_effects_df, aes(x = Country, y = predicted_prob)) + 
+  geom_bar(stat = "identity", fill = "grey", color = "black") +
+  labs(
+       x = "Country",
+       y = "Predicted Probability") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  # Rotate country names for readability
+  theme_minimal()
+```
+
+\newpage
 # References

### 2024-12-02 scripts/03-clean_data.R
update scripts

@@ -37,6 +37,27 @@ cleaned_pac <- data %>%
 cleaned_pac
 
 
+combined_data <- combined_data %>%
+  mutate(across(`1925`:`2023`, as.numeric))
+koreadata <- koreadata %>%
+  mutate(across(`1925`:`2023`, as.numeric))
+
+combined_data <- combined_data  %>%
+  select(-`Whole Country/Province/State`, -`Data Type`, -`Reporting Area`, -Species, -`Catch Type`)
+koreadata <- koreadata  %>%
+  select(-`Whole Country/Province/State`, -`Data Type`, -`Reporting Area`, -Species, -`Catch Type`)
+
+long_data_combined <- combined_data %>%
+  pivot_longer(
+    cols = starts_with("19") | starts_with("20"),  # Select columns starting with "19" or "20"
+    names_to = "Year",
+    values_to = "Catch"
+  ) %>%
+  mutate(Year = as.numeric(Year))
+
+write_csv(long_data_combined, "./data/02-analysis_data/clean_long_data_combined.csv")
+arrow::write_parquet(long_data_combined, "./data/02-analysis_data/clean_long_data_combined.csv")
+
 #### Save data ####
 write_csv(cleaned_pac, "./data/02-analysis_data/cleaned_pacific_allyear.csv")
 arrow::write_parquet(cleaned_pac, "./data/02-analysis_data/cleaned_pacific_allyear.parquet")

### 2024-12-02 scripts/06-model_data.R

@@ -77,5 +77,3 @@ saveRDS(
 
 
 
-
-

### 2024-12-02 paper/paper.qmd
last paper

@@
 #| message: false
-data <- read.xlsx('../data/01-raw_data/NPAFC_Catch_Stat-1925-2023.xlsx', sheet = 1) 
+data <- read.xlsx("../data/01-raw_data/NPAFC_Catch_Stat-1925-2023.xlsx", sheet = 1)
 colnames(data) <- data[1, ]
@@
 
-cleaned_pac2 <-cleaned_pac2 %>%
+cleaned_pac2 <- cleaned_pac2 %>%
   select(-Species)
@@
 combined_data <- bind_rows(cleaned_pac, koreadata)
-
 ```
@@
 
+write_csv(long_data_combined, "../data/02-analysis_data/clean_long_data_combined.csv")
+
 long_data_korea <- koreadata %>%
@@
 
-Estimand is the number of fishes that were caught by each country, for every year. Or more specifically what the rate was and if they are more likely to be fished for commercial purposes.
+Overfishing and its ecological consequences represent critical challenges for sustainable resource management.
+In marine ecosystems, excessive fishing reduces fish stocks and destabilizing food chains.
+These effects are especially present in anadromous fish populations in the North Pacific Ocean, which plays an important role in both local ecosystems and global seafood markets.
+Despite the international regulations and monitoring efforts, disparities in fishing practices and compliance with catch limits persist.
+Understanding the interactions between fishing rates, international contributions, and sustainability is this papers main goal.
+
+This paper investigates fishing dynamics among the five key members of the North Pacific Anadromous Fish Commission (NPFAC): the United States, Japan, Korea, Canada, and Russia.
+Our primary focus is on two aspects: (1) fluctuations in annual fish catch rates from 1925 to 2024, and (2) the probability of countries engaging in commercial fishing.
+We use Bayesian Generalized Linear Mixture Models, to figure out the rate of annual fish catch per year, where we find that every year the number fish caught increases by 721,000 fish.
+These models reveals that US and Russia has a high rate of change than Japan, Korea, Canada but collectively they are a positive rate of change.
+Our estimand is the number of fishes that were caught by each country, for every year.
+Or more specifically what the rate was and if they are more likely to be fished for commercial purposes.
+
+We are trying to bridge the gap between the data that contains the number of fish caught and the sustainability concerns about said numbers.
+This research has been done already, most likely, however, we would like to introduce statistics and methods of observational data to aid inn this research.
+Although NPFAC aims to limit fishing activity, the catch data often suggest otherwise, especially when countries seem to fish more than they were advised to.
+For instance, the U.S.
+appears to exceed its allowable catch as seen in the discussion section later.
+Furthermore, while Russia emerges as a dominant provider of anadromous fish in our data, global perspectives point to China as the leading fishing nation [@ritchie2023fish], which is a country that is not in our data.
+
+Through our analysis, we identify critical patterns: fluctuations in fish stocks align with overfishing impacts, and fishing for commercial purposes dominates in all observed countries.
+The findings have significant implications for ecological preservation and the global seafood supply chain, as they underscore the importance of stricter compliance and adaptive management in maintaining fish populations.
+The paper is structured as follows: the data and methodologies are introduced in Section 2; results, including model summaries and predictions, are presented in Section 3; a discussion of sustainability and dominant fishing practices follows in Section 4; finally, conclusions and policy recommendations are outlined in Section 5.
+There are 3 appendicies: one for methods in survey, sampling and observational data, one for datasheet to explain the data, and the last as a model card exaplining that models.
 
@@
 This paper look at multiple variables. We will go through them one by one:
+
 - First, variable we look at is Country which are either "Canada", "Russia", "Korea", "Japan", and "United States", all representing the countries that are members of this organization.
 This data was left unchanged.
+
 - Next variables we look at is "Whole Country/Province/State", where the instances of these variables may either be Whole country, or the different states or provinces that was fishing and gathered that data.
@@
 This was because we were more interested in comparing the fishing trends between countries rather then within a country.
+
 - The next variable is "Reporting Area" which accounts for where the fishes were caught.
@@
 Although there is a lot of interesting information to uncover if we did a deeper analysis on each fish, but, we decided that the best way to compare the fishing trends between countries would be to just look at the total fishes caught.
+
 - The next variable we used and actually analyse is the "catch type".
 This tells us whether is the fishes were caught for commercial purposes (caught for profit purposes like selling), sporting purposes (which means they were caught recreationaly) or subsistence purposes (which means they were caught to provide food, not as a profit).
+
 - The last variable we filtered was "Data Type" which was the unit that these numbers were reported in.
@@
   geom_point(size = 2, alpha = 0.8) + # Points with smaller size
-  labs(x = "Year", y = "Catch") +
+  labs(x = "Year", y = "Catch (Thousands of Fish)") +
   scale_y_continuous(labels = scales::comma) + # Format y-axis with commas
@@
   theme(legend.title = element_blank())
-
 ```
@@
   geom_point(size = 2, alpha = 0.8) + # Points with smaller size
-  labs( x = "Year", y = "Catch") +
+  labs(x = "Year", y = "Catch (Thousands of Fish)") +
   scale_y_continuous(labels = scales::comma) + # Format y-axis with commas
@@
     vjust = -0.1, # Adjust vertical position of text
-            size = 2) +  # Smaller font size for the numbers
+    size = 2
+  ) + # Smaller font size for the numbers
   labs(
     x = "Catch Type",
-       y = "Catch Data") +
+    y = "Catch Data (Thousands of Fish)"
+  ) +
   scale_y_continuous(labels = scales::comma) +
@@
   theme(legend.title = element_blank())
-
 ```
 @fig-total_bar is a great way to analyse the different reasons why these fishes are caught.
-We narrowed the year to 2022 as we predict the fishing industry has recovered after covid and will show accurate results.
+We narrowed the year to 2022 as we predict the fishing industry has recovered after COVID and will show accurate results.
 Here we see that most fishes are caught due to commercial purposes where they sell the fish rather then for sporting or subsistence purposes.
@@
 
-#model_formula <- Catch ~ Year + (1 | Country)
+# model_formula <- Catch ~ Year + (1 | Country)
 
-#bayesian_model <- stan_glmer(
+# bayesian_model <- stan_glmer(
 #  formula = model_formula,
@@
 #  cores = 4
-#)
-#saveRDS(
+# )
+# saveRDS(
 #  bayesian_model,
 #  file = "../models/bayesian_model1.rds"
-#)
+# )
 bayesian_model <-
   readRDS(file = here::here("models/bayesian_model1.rds"))
+```
+
+### Sub Model
+This small model show cases the different rates of fishes caught per year for each country. It was a simple linear regression:
+
+$Catch = \beta_0 + \beta_1 Year$
+
+For each country. We will only look at the $\beta_1$ value in @tbl-eachCountry.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| label: tbl-eachCountry
+#| tbl-cap: "A Simple Linear Regression for Each Country"
+# Load necessary libraries
+library(dplyr)
+library(knitr)
+
+# Fit a linear model for each country and extract the slope (rate of change)
+rate_of_change <- long_data_combined %>%
+  group_by(Country) %>%
+  do(model = lm(Catch ~ Year, data = .)) %>% # Fit linear model per country
+  summarise(
+    country = Country,
+    rate_of_change = round(coef(model)[2], 2)
+  ) # Extract the slope and round to 2 decimals
+
+# Calculate the mean rate of change
+mean_rate_of_change <- round(mean(rate_of_change$rate_of_change), 2)
+
+# Add the mean rate of change as a new row
+rate_of_change_with_mean <- bind_rows(
+  rate_of_change,
+  tibble(country = "Mean", rate_of_change = mean_rate_of_change)
+)
+
+# Display the result using kable for a nicely formatted table
+kable(rate_of_change_with_mean, caption = "Rate of Change for Each Country (Including Mean)", digits = 2)
 ```
@@
 
-We predict that the outcome or rather the probability that the fishes caught were commercially would be positive.
+We predict that the outcome or rather the probability that the fishes caught were commercially would be very high; we are expecting more than 50%.
 This is because as we were looking at the graphs in the data section, we notice that @fig-total_bar shows most of the fishes caught were commercially and then as a sport and then as a subsistence.
@@
 # Specify priors
-#priors <- c(
+# priors <- c(
 #  set_prior("normal(0, 2.5)", class = "Intercept"),   # Prior for the intercept
 #  set_prior("normal(0, 2.5)", class = "sd")           # Prior for the standard deviation of random effects (Country)
-#)
+# )
 
-#cleaned_pac2$Is_Commercial <- ifelse(cleaned_pac2$`Catch Type` == "Commercial", 1, 0)
+# cleaned_pac2$Is_Commercial <- ifelse(cleaned_pac2$`Catch Type` == "Commercial", 1, 0)
 
 # Fit the model using brms
-#bayesian_model_commercial_brms <- brm(
+# bayesian_model_commercial_brms <- brm(
 #  formula = Is_Commercial ~ (1 | Country),  # Random effect for Country
@@
 # seed = 123
-#)
+# )
 
-#saveRDS(
+# saveRDS(
 #  bayesian_model_commercial_brms,
 #  file = "../models/bayesian_model2.rds"
-#)
+# )
 
@@
 #| results: hide
-##| label: fig-randomInter
-##| fig-cap: "Distribution of Random Intercepts for Country"
+## | label: fig-randomInter
+## | fig-cap: "Distribution of Random Intercepts for Country"
 # Extract random effects for Country
@@
 # Now, plot the distribution of the random intercepts for Country
-#library(ggplot2)
+# library(ggplot2)
 
-#ggplot(random_effects_df, aes(x = Estimate.Intercept)) + 
+# ggplot(random_effects_df, aes(x = Estimate.Intercept)) +
 #  geom_histogram(binwidth = 0.1, fill = "grey", color = "black", alpha = 0.7) +
@@
 #  theme_minimal()
-
 ```
@@
     x = "Country",
-       y = "Predicted Probability") +
+    y = "Predicted Probability"
+  ) +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) + # Rotate country names for readability
   theme_minimal()
-
 ```
@@
 # Discussion {#sec-disc}
+We have shown a lot of graphs and the results from the graphs but now we will give context to the results and see where they come from.
 
-## Story of the Data
+## What Is Done In This Paper? (A Summary)
+As we discussed in the data section, we have fluctuations in our @fig-long_comb and @fig-korea data because of the idea that when you fish too much, you leave less fish to repopulate the ecosystem and hence there is less fish in general.
+This causes the next couple years to have less fish in the ocean and hence less fish to hunt.
+This is what we investigate in the first model, @tbl-modelSum-2.
+Next we take a look at @fig-total_bar where we see that Russia is leading in the commercial fishing catagory for 2022.
+We also see that out of the 5 countries, all of them commercially fish more than they fish as a sport or subsistence.
+In fact Forbes Russia, documents the largest fishing organisations of Russia, highlighting their large fishing industry [@RussiaFishMajor].
+The U.S.
+and Japan being second and third makes sense as U.S.
+is known for their booming economy due to them sharing the Pacific Ocean's coast with Japan.
+We predict the probability that a country is commercially fishing in the year 2022, in @fig-predProb.
+The country that catches the most fish is also shown in @ritchie2023fish.
 
-## Weaknesses and next steps
+The more interesting part of the paper is the models we created.
+As mentioned earlier we created a model to see how many fishes were caught per year using a Bayesian Generalized Linear Mixture Model.
+We show the model summary in @tbl-modelSum-2, where we see that the intercept is at -1,367,355 which tells us that for a "baseline" country, at year 0, there were negative fishes caught.
+This logically doesn't make sense however.
+The reason why this number was outputted was because of the fact that all 5 countries have varying rates between how much fish will be caught per year, with some values being negative and some positive, some being very steep and others not.
+This is shown in @tbl-eachCountry.
+We see that the average of all of the rates in @tbl-eachCountry is 606 thousand fishes per year which is close to the 721.74 thousand fish per year that we get.
+The slight difference is because in @tbl-modelSum-2 we consider the year as well as the random effect of the country.
+Any how this summary of @tbl-modelSum-2 and @tbl-eachCountry perfectly aligns with the numbers we see in @fig-long_comb where we see that the rates are rising for Russia and U.S while decreasing for Japan, Korea and Canada.
+However, U.S and Russia have a higher rate, 2347 thousand fish per year and 1448 thousand fish per year respectively, which causes the overall positive rate in @tbl-modelSum-2.
+Regarding our prediction in @sec-mod1's model justification, we were correct as the rate of fish caught per year is indeed positive.
+this could be easily predicted from @fig-long_comb.
+
+Another model we explored was what is the probability that a country fished for commercial purposes in 2022.
+This was shown in the model summary, @tbl-modelSum-1.
+This showed that the intercept is 0.84, which says that there is 69.8% chance that a "baseline" country has fished for commercial purposes in 2022.
+This goes with our prediction that the probability will be very high in @sec-mod2's model justification section.
+This makes sense as most countries do not fish tens of millions of fish just as food for the country.
+Especially the countries that are part of NPFAC.
+They are top providers of seafood to the global fish demand and hence it makes sense that they would fish commercially for the most part.
+
+## What Does This Mean?
+There are both positive and negative conclusions that can be drawn from this data.
+
+   The first biggest, and negative, point is the issues with sustainability and overfishing.
+@WPRFMC2024 shows us the catch limit of 3 out of the 5 NPFAC members; It shows the US, Japan and Korea.
+Japan has a catch limit of 18,000 metric ton of fish, Korea has 13,000 metric tonne, and U.S has 4,000 metric tonne (recorded for the year 2018 to 2021).
+On average 1 metric tonne of fish is around 11,000 fish so according to that Japan, Korea and U.S are allowed 198000000, 143000000, and 33000000 fish per year.
+A quick look at our data showed that there is 35173190, 95993, 118708471 fish caught in 2018.
+This means that Japan and Korea was the only one within that threshold, while U.S.
+exceeded it by a lot.
+We must also account for the demand of fish as well.
+the population of U.S is much larger then Korea and Japan, and the States must also supply the fish to other countries as well.
+@ritchie2023fish shows that, that catch limit can be adjusted according to the fish stock (amount of fish available) and the maximum amount you can catch without loosing on food and income.
+Although we keep this in account a difference of 85 million is too much for the ecosystem to handle.
+The reason NPFAC was formed was to keep these numbers under control.
+And this is only considering the U.S.
+and not Russia, the other fishing giant in the North Pacific.
+This was one of the main reasons why this paper was written.
+The anadromous fish population in the Pacific is important as they keep both the lower level of the food chain under control (such as smaller fish) while also providing food to the higher levels of the food chain (larger fish such as sharks).
+
+   The second aim of this paper is to see the dominant provider of the fish industry.
+Our data shows that Russia is by far the largest anadromous fish provider in the North Pacific but, @ritchie2023fish shows different as, it shows China is the biggest.
+The reason will be shown in @sec-weak.
+With that aside out of all the members of the NPFAC's we see that U.S.
+and Russia supply the most anadormous fish which is supported by @ritchie2023fish.
+
+## What Did We Miss? {#sec-weak}
+We struggled with the data as many of the values we not available.
+This makes sense as the data starts at 1925 and not everything has been recorded from 100 years.
 Korea didn't report data until 1969 so they are left out of the first model in general.
+
+We are also missing a lot of countries that fish in the pacific such as Indonesia, and China who fish in the pacific as well.
+
+Our next steps are to gather data from these countries as well and the same analysis on it to see if our results change.
+We would be looking at the same variables such as the rate of fish caught per year and the probability that a country was fishing commercially.
+Again we expect to see a positive rate and a high probability that a country was fishing commercially.
 
@@
   geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
-  labs(title = "Model Predictions by Country with 95% Credible Intervals", 
+  labs(
+    title = "Model Predictions by Country with 95% Credible Intervals",
     x = "Country",
-       y = "Mean Prediction") +
+    y = "Mean Prediction"
+  ) +
   theme_minimal()
@@
   geom_line() +
-  facet_wrap(~ Country, scales = "free_y") +  # Facet by Country
+  facet_wrap(~Country, scales = "free_y") + # Facet by Country
   labs(title = "Predictions of Catch by Year and Country", x = "Year", y = "Mean Prediction") +
@@
     x = "Country",
-       y = "Predicted Probability") +
+    y = "Predicted Probability"
+  ) +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) + # Rotate country names for readability

### 2024-12-02 scripts/03-clean_data.R
script update

@@ -56,7 +56,7 @@ long_data_combined <- combined_data %>%
   mutate(Year = as.numeric(Year))
 
 write_csv(long_data_combined, "./data/02-analysis_data/clean_long_data_combined.csv")
-arrow::write_parquet(long_data_combined, "./data/02-analysis_data/clean_long_data_combined.csv")
+arrow::write_parquet(long_data_combined, "./data/02-analysis_data/clean_long_data_combined.parquet")
 
 #### Save data ####
 write_csv(cleaned_pac, "./data/02-analysis_data/cleaned_pacific_allyear.csv")
